Symbolic Weighted Language Models and\\ Parsing over Infinite Alphabets

### 2021-05-17 introduction.tex
example

% !TEX root = main.tex
%
% Introduction
%
Parsing %variant of membership problem for formal languages
%- given a language model and an input word $s$, compute a derivation of the model that yields $s$.
is the problem %process 
of structuring a linear representation in input %typically
(a finite word over an alphabet) according to a language model. % (a formal grammar).
% natural language, programming language, 
%
Most of the context-free parsing approaches~\cite{GruneJacobs08parsing}  
assume a finite and reasonably small input alphabet. %models and algorithms 
Such a restriction makes perfectly sense in the context of NLP tasks 
like constituency parsing or of programming languages compilers or interpreters.
Considering large or infinite alphabets can however be of practical interest
in other cases.
%
For instance, when dealing with large characters encodings such as UTF-16, % processing strings in 
\eg for vulnerability detection in Web-applications~\cite{dAntoni21CACM}, 
%
for the analyse (\eg validation or filtering) 
of data streams or serialization of structured documents 
(which may contain textual or numerical attributes)~\cite{Segoufin06csl}, 
or for processing timed execution traces~\cite{Bouyer03algebraic}.
% algebraic definition of a class of data languages 
% (notion of monoid recognizability, based on registers, comparable to Bojancszik et al. data words)
Regarding the latter case, we describe %briefly 
at the end of the paper 
a parse-based approach to automated music transcription,
where a symbolic music performance, 
presented as a sequence of timed musical events, % timestamped
% a symbolic representation of a music performance 
is converted into a structured score in Common Western Music Notation~\cite{foscarin:hal-01988990}.
%generalizes weighted parsing: 
%finding the best derivation for a weighted grammar.

@@

The main contributions of the paper are: 
($i$)~the introduction of automata, \SWA, transducers, \SWT (Section~\ref{section:SWA}),
and visibly pushdown automata \SWVPA (Section~\ref{sec:SWVPA}),
generalizing the corresponding classes of symbolic and weighted models, 
%for the weighted computation on (nested) words over infinite alphabets;  
($ii$)~a polynomial best-search algorithm for \SWVPA, %(Section~\ref{sec:best})
% a framework for parsing over infinite alphabets, 
and ($iii$)~a uniform framework (Section~\ref{sec:parsing}) for parsing over infinite alphabets, 
the keys to which are 
($iii.a$)~the \SWT-based definition of generic edit distances between input and output words,
and ($iii.b$)~the use of nested words, and \SWVPA, 
instead of syntax trees and grammars. %(\S~\ref{sec:trees}).
%
Finally, Section~\ref{sec:transcription} describes
the implemented application %of this approach 
to automated music transcription
that motivated this study.

>>>>

% !TEX root = main.tex
%
% Introduction
%
Parsing %variant of membership problem for formal languages
%- given a language model and an input word $s$, compute a derivation of the model that yields $s$.
is the problem %process 
of structuring a linear representation in input %typically
(a finite word over an alphabet) according to a language model. % (a formal grammar).
% natural language, programming language, 
%
Most of the context-free parsing approaches~\cite{GruneJacobs08parsing}  
assume a finite and reasonably small input alphabet. %models and algorithms 
Such a restriction makes perfectly sense in the context of NLP tasks 
like constituency parsing or of programming languages compilers or interpreters.
Considering large or infinite alphabets can however be of practical interest
in other cases.
%
For instance, when dealing with large characters encodings such as UTF-16, % processing strings in 
\eg for vulnerability detection in Web-applications~\cite{dAntoni21CACM}, 
%
for the analyse (\eg validation or filtering) 
of data streams or serialization of structured documents 
(which may contain textual or numerical attributes)~\cite{Segoufin06csl}, 
or for processing timed execution traces~\cite{Bouyer03algebraic}.
% algebraic definition of a class of data languages 
% (notion of monoid recognizability, based on registers, comparable to Bojancszik et al. data words)
Regarding the latter case, we briefly mention
at the end of the paper 
a parse-based approach to automated music transcription~\cite{foscarin:hal-01988990}
that motivated the present work. 
In this problem, a symbolic music performance, 
presented as a sequence of timed musical events, % timestamped
% a symbolic representation of a music performance 
is converted into a structured score in Common Western Music Notation.
%generalizes weighted parsing: 
%finding the best derivation for a weighted grammar.

@@

The main contributions of the paper are: 
($i$)~the introduction of automata, \SWA, transducers, \SWT (Section~\ref{section:SWA}),
and visibly pushdown automata \SWVPA (Section~\ref{sec:SWVPA}),
generalizing the corresponding classes of symbolic and weighted models, 
%for the weighted computation on (nested) words over infinite alphabets;  
($ii$)~a polynomial best-search algorithm for \SWVPA, %(Section~\ref{sec:best})
% a framework for parsing over infinite alphabets, 
and ($iii$)~a uniform framework (Section~\ref{sec:parsing}) for parsing over infinite alphabets, 
the keys to which are 
($iii.a$)~the \SWT-based definition of generic edit distances between input and output words,
and ($iii.b$)~the use of nested words, and \SWVPA, 
instead of syntax trees and grammars. %(\S~\ref{sec:trees}).
%
%Finally, Section~\ref{sec:transcription} describes
%the implemented application %of this approach 
%to automated music transcription
%that motivated this study.
